### 2018-04-19 Maths Notes/Further Pure/FP2/Overall Notes/FP2 Notes.tex
Summary Notes Update

\documentclass{article}[18pt]
\usepackage{../../../../format}
\lhead{A Level Maths - FP2}

\begin{document}
\begin{center}
\underline{\huge FP2}
\end{center}
\section{First order differential equations}
\subsection{Solving first order DE using an integrating factor}
Solving $\dfrac{dy}{dx}+P(x)y=Q(x)$\\
\\
IF(Integrating factor) is found by finding $\mathlarger{e^{\int p(x) \ dx}}$ And multiplying the DE by the IF.\\
\\
This will result in the DE being in the form:
$$f(x)\frac{dy}{dx}+f'(x)y$$
This form can then be shortened by integrating:
$$\int f'(x)g(x)+f(x)g'(x) dx=f(x)g(x)+c$$
Integrate both sides then simplify
\section{Further complex numbers}
\subsection{Converting between forms}
When converting from $x+iy$ to a form in r and $\theta$, take the angle from the positive $x$ axis
\subsection{Multiplying and dividing complex numbers}
It is easiest to use the exponential form, then convert if needed
\subsubsection{Multiplying}
$$Z_1Z_2=r_1e^{i\theta_1}\times r_2e^{i\theta_2}=r_1r_2e^{i(\theta_1+\theta_2)}$$
\subsubsection{Dividing}
$$\frac{Z_1}{Z_2}=r_1e^{i\theta_1}\div r_2e^{i\theta_2}=\frac{r+1}{r+2} e^{i(\theta_1-\theta_2)}$$
\subsection{De Moivre's theorem}
This is given on the data sheet
\subsubsection{Z formulas}
If $z=\cos\theta+i\sin\theta$
$$z+\frac{1}{z}=2\cos\theta$$
$$z-\frac{1}{z}=2i\sin\theta$$
$$z^n+\frac{1}{z^n}=2\cos n\theta$$
$$z^n-\frac{1}{z^n}=2i\sin n\theta$$
\subsection{Loci on the complex plane}
\begin{tabularx}{\textwidth}{|X|X|}
\hline
Equation&Description\\
\hline
{\Large$|z-z_1|=r$}&A circle centre $(x_1,y_1)$ with a radius r\\
\hline
{\Large$|z-z_1|=|z-z_2|$}&A perpendicular bisector of the line segment joining points $z_1$ and $z_2$\\
\hline
{\Large$\arg(z-z_1)=\theta$}&The half line from a fixed point $z_1$, making an angle $\theta$ with the positive real axis\\
\hline
{\Large$\arg(\frac{z-z_1}{z-z_2})=\theta$}&An arc between the points $z_1$ and $z_2$ where the angle the lines from $z_1$ and $z_2$ to any point on the arc is $\theta$\\
\hline
\end{tabularx}
\subsection{Translations}
\begin{itemize}
\item $w=z+a+ib$ represents a translation with translation vector $\begin{pmatrix}
a\\b
\end{pmatrix}$
\item $w=kz$ represents an enlargement with scale factor $k$ centre $(0,0)$
\item $w=kz+a+ib$ represents an enlargement scale factor $k$ centre $(0,0)$ followed by a translation with translation vector $\begin{pmatrix}
a\\b
\end{pmatrix}$
\item $w=z^2$ multiply a shape by itself, for example a circle of radius 4 would go to radius 16
\end{itemize}
\section{Inequalities}
We can build upon our previous algebraic skills in order to solve more complex inequalities\\
Remember:
\begin{itemize}
\item Don't multiply anything that could be negative - use "squared" things
\item Find the critical values (f(x)=0)
\item Sketch the graph to solve
\end{itemize}
\section{Maclaurin and Taylor Series}
Use the formulas on the data sheet
\subsection{Solving differential equations using the Taylor expansion}
From the differential equation, calculate the values of $\frac{dy}{dx}$, $\frac{d^2y}{dx^2}$ etc up to whatever is needed.\\
Then substitute those values into the Taylor series to solve the differential equation

>>>>

\documentclass{article}[18pt]
\usepackage{../../../../format}
\lhead{A Level Maths - FP2}
\usepackage{tabularx}
\begin{document}
\begin{center}
\underline{\huge FP2}
\end{center}
\section{First order differential equations}
\subsection{Solving first order DE using an integrating factor}
Solving $\dfrac{dy}{dx}+P(x)y=Q(x)$\\
\\
IF(Integrating factor) is found by finding $\mathlarger{e^{\int p(x) \ dx}}$ And multiplying the DE by the IF.\\
\\
This will result in the DE being in the form:
$$f(x)\frac{dy}{dx}+f'(x)y$$
This form can then be shortened by integrating:
$$\int f'(x)g(x)+f(x)g'(x) dx=f(x)g(x)+c$$
Integrate both sides then simplify
\section{Further complex numbers}
\subsection{Converting between forms}
When converting from $x+iy$ to a form in r and $\theta$, take the angle from the positive $x$ axis
\subsection{Multiplying and dividing complex numbers}
It is easiest to use the exponential form, then convert if needed
\subsubsection{Multiplying}
$$Z_1Z_2=r_1e^{i\theta_1}\times r_2e^{i\theta_2}=r_1r_2e^{i(\theta_1+\theta_2)}$$
\subsubsection{Dividing}
$$\frac{Z_1}{Z_2}=r_1e^{i\theta_1}\div r_2e^{i\theta_2}=\frac{r+1}{r+2} e^{i(\theta_1-\theta_2)}$$
\subsection{De Moivre's theorem}
This is given on the data sheet
\subsubsection{Z formulas}
If $z=\cos\theta+i\sin\theta$
$$z+\frac{1}{z}=2\cos\theta$$
$$z-\frac{1}{z}=2i\sin\theta$$
$$z^n+\frac{1}{z^n}=2\cos n\theta$$
$$z^n-\frac{1}{z^n}=2i\sin n\theta$$
\subsection{Loci on the complex plane}
\begin{tabularx}{\textwidth}{|X|X|}
\hline
Equation&Description\\
\hline
{\Large$|z-z_1|=r$}&A circle centre $(x_1,y_1)$ with a radius r\\
\hline
{\Large$|z-z_1|=|z-z_2|$}&A perpendicular bisector of the line segment joining points $z_1$ and $z_2$\\
\hline
{\Large$\arg(z-z_1)=\theta$}&The half line from a fixed point $z_1$, making an angle $\theta$ with the positive real axis\\
\hline
{\Large$\arg(\frac{z-z_1}{z-z_2})=\theta$}&An arc between the points $z_1$ and $z_2$ where the angle the lines from $z_1$ and $z_2$ to any point on the arc is $\theta$\\
\hline
\end{tabularx}
\subsection{Translations}
\begin{itemize}
\item $w=z+a+ib$ represents a translation with translation vector $\begin{pmatrix}
a\\b
\end{pmatrix}$
\item $w=kz$ represents an enlargement with scale factor $k$ centre $(0,0)$
\item $w=kz+a+ib$ represents an enlargement scale factor $k$ centre $(0,0)$ followed by a translation with translation vector $\begin{pmatrix}
a\\b
\end{pmatrix}$
\item $w=z^2$ multiply a shape by itself, for example a circle of radius 4 would go to radius 16
\end{itemize}
\section{Inequalities}
We can build upon our previous algebraic skills in order to solve more complex inequalities\\
Remember:
\begin{itemize}
\item Don't multiply anything that could be negative - use "squared" things
\item Find the critical values (f(x)=0)
\item Sketch the graph to solve
\end{itemize}
\section{Maclaurin and Taylor Series}
Use the formulas on the data sheet
\subsection{Solving differential equations using the Taylor expansion}
From the differential equation, calculate the values of $\frac{dy}{dx}$, $\frac{d^2y}{dx^2}$ etc up to whatever is needed.\\
Then substitute those values into the Taylor series to solve the differential equation
\section{Polar Coordinates}
\begin{tabularx}{\textwidth}{|X|X|}
\hline
Equation&Description\\
\hline
$r=a$&A circle, centre (0,0) with a radius of a\\
\hline
$\theta=\alpha$&A half line starting from (0,0) making an angle $\alpha$ with the initial line\\
\hline
$r=a\theta$&A spiral starting at the origin\\
\hline
$r=a(1+\cos\theta)$&A cardioid with x intercept a\\
\hline
$r^2=a^2\cos2\theta$&An infinity symbol with x intercept of a\\
\hline
$r^2=a^2\sin2\theta$&Infinity symbol rotated by $\frac{\pi}{4}$\\
\hline
$r=a\sin3\theta$&A windmill with 3 blades\\
\hline
\end{tabularx}
\newpage
\subsection{Integration}
$$\textrm{Area}=\frac{1}{2}\int_\alpha^\beta r^2 \ d\theta$$
The area of the sector is bounded by the half lines $\theta=\alpha$ and $\theta=\beta$
\subsection{Differentiation}
To find tangents parallel and perpendicular to the initial line:\\
\\
Parallel: $\frac{dy}{d\theta}=0$\\
Perpendicular: $\frac{dx}{d\theta}=0$
\section{Second order differential equations}
{\renewcommand{\arraystretch}{2}
\begin{tabularx}{\textwidth}{|X|X|}
\hline
\textbf{Solution to auxiliary equation}&\textbf{General solution}\\
\hline
Two real roots $b^2>4ac$&$Ae^{\alpha x}+Be^{\beta x}$\\
\hline
One real root $b^2=4ac$&$(A+bx)e^{\alpha x}$\\
\hline
Imaginary only root $b^2<4ac$&$y=Ae^{\alpha x}+Be^{-\alpha x}$\\
\hline
Complex root $b^2<4ac$&$y=Ae^{(\beta+\alpha i)x}+Be^{(\beta-\alpha i)x}$\\
\hline
\end{tabularx}}
\subsection{Second ODE=f(x)}
\subsubsection{Standard forms of f(x)}
$f(x)=\lambda$\\
$f(x)=\lambda+\mu x$\\
$f(x)=\lambda+\mu x+\nu x^2$\\
$f(x)=ke^{px}$\\
$f(x)=m\cos\omega x$\\
$f(x)=m\sin\omega x$\\
$f(x)=m\cos\omega x\pm n\sin\omega x$
\\
To solve these find $\frac{dy}{dx}$ and $\frac{d^2y}{dx^2}$ and substitute them into the Second ODE, solving for the unknowns, then add them to the general solution.\\
\\
If there is a clash of terms between the CF and PI, then an alternate PI will be given to you.
\subsection{Boundary conditions}
If boundary conditions are given to you, the values of the unknowns can be found. Substitute the boundary conditions in and rearrange to find unknowns.
\newpage
\subsection{Substitution}
Transformations can be used to simplify differential equations, leaving $\frac{d^2y}{dx^2}$ with no coefficient.\\
Method:\\
\textcolor{red}{\large Find $\frac{dy}{dx}$ in terms of $\frac{dy}{du}$}
$$x=e^u \quad \frac{dx}{du}=e^u \quad \frac{du}{dx}=e^{-u} \quad 
\frac{dy}{dx}=\frac{dy}{du}\times\frac{du}{dx}=\mathbf{e^{-u}\frac{dy}{du}}
$$
\textcolor{red}{\large Find $\frac{d^2y}{dx^2}$ in terms of $\frac{d^2y}{du^2}$ and $\frac{dy}{du}$}\\
\textcolor{red}{Find $\frac{dy}{du}\Big(\frac{dy}{dx}\Big)$ by applying the product rule}
$$\frac{dy}{du}\Big(\frac{dy}{dx}\Big)=-e^{-u}\frac{dy}{du}+e^{-u}\frac{d^2y}{du^2}$$
\textcolor{red}{Multiply through by $\frac{du}{dx}$ to find the value of $\frac{d^2y}{dx^2}$}
$$\frac{d^2y}{dx^2}=\frac{dy}{du}\Big(\frac{dy}{dx}\Big)\times\frac{du}{dx}$$
$$\frac{d^2y}{dx^2}=e^{-u}\Big(-e^{-u}\frac{dy}{du}+e^{-u}\frac{d^2y}{du^2}\Big)$$
$$\frac{d^2y}{dx^2}=\mathbf{e^{-2u}\Big(-\frac{dy}{du}+\frac{d^2y}{du^2}\Big)}$$
\section{Series}
Summations can sometimes be simplified using the method of differences\\
\\
This involves a summation where one term is subtracted from another, often gained from partial fractions. The terms in the middle of the series will then cancel with each other, leaving the summation just the terms at the beginning at the end, by summing these a simpler form can be gained.

### 2018-04-19 Maths Notes/Mechanics/M2/Centres of mass/Centres of mass.tex
Update of all files to shorten preamble
Tidied to work with all linux systems
Summary Notes Update

\documentclass{article}[18pt]
\usepackage[utf8]{inputenc}
\usepackage[margin=0.7in]{geometry}
\usepackage{parselines} 
\usepackage{amsmath}
\usepackage{titlesec}
\usepackage{pgfplots}
\usepackage{graphicx}
\usepackage[english]{babel}
\usepackage{fancyhdr}
\usepackage{gensymb}
\usepackage{tabularx}

\pgfplotsset{width=10cm,compat=1.9}

\titlespacing\section{0pt}{14pt plus 4pt minus 2pt}{0pt plus 2pt minus 2pt}
\newlength\tindent
\setlength{\tindent}{\parindent}
\setlength{\parindent}{0pt}
\renewcommand{\indent}{\hspace*{\tindent}}

\pagestyle{fancy}
\fancyhf{}
\rhead{Sam Robbins 13SE}
\lhead{A Level Maths - M2}
\rfoot{Page \thepage}

\begin{document}
\begin{center}
\underline{\huge Centres of mass}
\end{center}
\begin{obeylines}
\section{Centre if mass of a discrete mass distribution}
$\bar{x}=\dfrac{\Sigma m_ix_i}{\Sigma m_i}$\\
$\bar{y}=\dfrac{\Sigma m_iy_i}{\Sigma m_i}$\\
\textbf{Example}\\
\end{obeylines}
\begin{tabular}{|c|c|c|c|}
 \hline
 Mass&$2$&$3$&$2$\\
 \hline
 $x$&2&3&-3\\
 \hline
 $y$&3&6&2\\
 \hline

\end{tabular}
\\
\\
\begin{obeylines}
$\bar{x}=\dfrac{2\times 2+3\times3+2\times-3}{2+3+2}=1$
$\bar{y}=\dfrac{2\times3+3\times6+2\times2}{2+3+2}=4$
\section{Uniform laminae}
For a triangular lamina the centre of mass is $\frac{2}{3}$ along the line from the vertex to the middle of the line opposite.
For a sector of a circle, radius r, where the angle at the centre is $2\alpha$ the centre of mass is $\dfrac{2r\sin\alpha}{3\alpha}$
\section{Rods}
In a circular arc, radius r, where the angle at the centre is $2\alpha$, the centre of mass is $\dfrac{r\sin\alpha}{\alpha}$ away from the centre.
\section{Equilibrium}
To avoid tipping, the lime of action of the weight must be within the side of the lamina in contact with the plane.
If a lamina is suspended from a fixed point, the centre of mass will be vertically below the point of suspension.
Assumptions made in equilibrium calculations:
\begin{itemize}
\item No friction at the point of suspension
\item The mass of each area is uniform
\item The mass is uniform at the join
\end{itemize}

@@

\textbf{1. Find the masses and locations of all areas or rods and put in a table}\\
\\
{\renewcommand{\arraystretch}{1.5}
\begin{tabularx}{\textwidth}{|X|X|X|}
\hline
Mass(m)&$4a^2$&$-a^2$\\
\hline
Distance From AF(x)&$a$&$\frac{3}{2}a$\\
\hline
mx&$4a^3$&$-\frac{3}{2}a^3$\\
\hline
\end{tabularx}}
\\
\\
\textbf{2. Use the formula to find the centre of mass}\\
\\
\textcolor{red}{$\bar{x}=\dfrac{\Sigma m_ix_i}{\Sigma m_i}$}\\
\\
$\bar{x}=\dfrac{\frac{5}{2}a^3}{3a^2}=\dfrac{5a}{6}$\\
\\
\textit{The lamina is freely suspended from A and hangs in equilibrium.}\\
\textit{(b) Find, in degrees to one decimal place, the angle which AF makes with the vertical.}\\
\\
\textbf{Use symmetry to find the y coordinate of the centre of mass}\\
\\
\begin{tikzpicture}[scale=1.9]
\tikzstyle{ann} = [fill=white,font=\footnotesize,inner sep=1pt]

\draw[black, thick] (0,0) -- (0,2) -- (1,2) -- (1,1) -- (2,1) -- (2,0) -- cycle;
\draw[arrows=<->](0,5/6)--(5/6,5/6);
\node[ann] at (5/12,5/6) {$\frac{5}{6}$};
\draw [blue,dashed] (0,0) -- (2,2);
\draw[arrows=<->](5/6,0)--(5/6,5/6);
\node[ann] at (5/6,5/12) {$\frac{5}{6}$};
\end{tikzpicture}
\\
\\
\textbf{Re-draw to find angle}\\
\begin{tikzpicture}[scale=1.9]
\tikzstyle{ann} = [fill=white,font=\footnotesize,inner sep=1pt]
\tikzstyle{angle} = [fill=none,font=\tiny,inner sep=1pt]
\draw[black, thick] (0,0) -- (0,2) -- (1,2) -- (1,1) -- (2,1) -- (2,0) -- cycle;
\draw[arrows=<->](0,5/6)--(5/6,5/6);
\node[ann] at (5/12,5/6) {$\frac{5}{6}$};
\draw[arrows=<->](5/6,5/6) -- (5/6,2);
\node[ann] at (5/6,17/12) {$\frac{7}{6}$};
\draw [blue,dashed] (0,2) -- (5/6,5/6);
\draw[thick] (5/24,41/24) arc (0:-90:0.2);
\node[ann] at (0.1,1.7) {$\theta$};
\end{tikzpicture}\\
\\
$\tan(\theta)=\dfrac{5/6}{7/6}$\\
\\
$\theta=35.5\degree$
\newpage

>>>>

\documentclass{article}[18pt]
\usepackage{../../../../format}
\lhead{A Level Maths - M2}

\begin{document}
\begin{center}
\underline{\huge Centres of mass}
\end{center}
\begin{obeylines}
\section{Centre of mass of a discrete mass distribution}
$\bar{x}=\dfrac{\Sigma m_ix_i}{\Sigma m_i}$\\
$\bar{y}=\dfrac{\Sigma m_iy_i}{\Sigma m_i}$\\
\textbf{Example}\\
\end{obeylines}
\begin{tabular}{|c|c|c|c|}
 \hline
 Mass&$2$&$3$&$2$\\
 \hline
 $x$&2&3&-3\\
 \hline
 $y$&3&6&2\\
 \hline

\end{tabular}
\\
\\
\begin{obeylines}
$\bar{x}=\dfrac{2\times 2+3\times3+2\times-3}{2+3+2}=1$
$\bar{y}=\dfrac{2\times3+3\times6+2\times2}{2+3+2}=4$
\section{Uniform laminae}
For a triangular lamina the centre of mass is $\frac{2}{3}$ along the line from the vertex to the middle of the line opposite.
For a sector of a circle, radius r, where the angle at the centre is $2\alpha$ the centre of mass is $\dfrac{2r\sin\alpha}{3\alpha}$
\section{Rods}
In a circular arc, radius r, where the angle at the centre is $2\alpha$, the centre of mass is $\dfrac{r\sin\alpha}{\alpha}$ away from the centre.
\section{Equilibrium}
To avoid tipping, the line of action of the weight must be within the side of the lamina in contact with the plane.
If a lamina is suspended from a fixed point, the centre of mass will be vertically below the point of suspension.
Assumptions made in equilibrium calculations:
\begin{itemize}
\item No friction at the point of suspension
\item The mass of each area is uniform
\item The mass is uniform at the join
\end{itemize}

@@

\textbf{1. Find the masses and locations of all areas or rods and put in a table}\\
\\
{\renewcommand{\arraystretch}{1.5}
\begin{tabularx}{\textwidth}{|X|X|X|}
\hline
Mass(m)&$4a^2$&$-a^2$\\
\hline
Distance From AF(x)&$a$&$\frac{3}{2}a$\\
\hline
mx&$4a^3$&$-\frac{3}{2}a^3$\\
\hline
\end{tabularx}}
\\
\\
\textbf{2. Use the formula to find the centre of mass}\\
\\
\textcolor{red}{$\bar{x}=\dfrac{\Sigma m_ix_i}{\Sigma m_i}$}\\
\\
$\bar{x}=\dfrac{\frac{5}{2}a^3}{3a^2}=\dfrac{5a}{6}$\\
\\
\textit{The lamina is freely suspended from A and hangs in equilibrium.}\\
\textit{(b) Find, in degrees to one decimal place, the angle which AF makes with the vertical.}\\
\\
\textbf{3. Use symmetry to find the y coordinate of the centre of mass}\\
\\
\begin{tikzpicture}[scale=1.9]
\tikzstyle{ann} = [fill=white,font=\footnotesize,inner sep=1pt]

\draw[black, thick] (0,0) -- (0,2) -- (1,2) -- (1,1) -- (2,1) -- (2,0) -- cycle;
\draw[arrows=<->](0,5/6)--(5/6,5/6);
\node[ann] at (5/12,5/6) {$\frac{5}{6}$};
\draw [blue,dashed] (0,0) -- (2,2);
\draw[arrows=<->](5/6,0)--(5/6,5/6);
\node[ann] at (5/6,5/12) {$\frac{5}{6}$};
\end{tikzpicture}
\\
\\
\textbf{4. Re-draw to find angle}\\
\begin{tikzpicture}[scale=1.9]
\tikzstyle{ann} = [fill=white,font=\footnotesize,inner sep=1pt]
\tikzstyle{angle} = [fill=none,font=\tiny,inner sep=1pt]
\draw[black, thick] (0,0) -- (0,2) -- (1,2) -- (1,1) -- (2,1) -- (2,0) -- cycle;
\draw[arrows=<->](0,5/6)--(5/6,5/6);
\node[ann] at (5/12,5/6) {$\frac{5}{6}$};
\draw[arrows=<->](5/6,5/6) -- (5/6,2);
\node[ann] at (5/6,17/12) {$\frac{7}{6}$};
\draw [blue,dashed] (0,2) -- (5/6,5/6);
\draw[thick] (5/24,41/24) arc (0:-90:0.2);
\node[ann] at (0.1,1.7) {$\theta$};
\end{tikzpicture}\\
\\
$\tan(\theta)=\dfrac{5/6}{7/6}$\\
\\
$\theta=35.5\degree$
\newpage

### 2018-04-19 Maths Notes/Mechanics/M2/Overall notes/M2 Notes.tex
Summary Notes Update

\begin{document}
\begin{center}
\underline{\huge Kinematics}
\end{center}
\begin{obeylines}
\section{Horizontal Projections}
For a \textbf{constant speed} use \textbf{$Speed=\dfrac{Distance}{Time}$}

For a \textbf{constant acceleration} use \textbf{SUVAT}

For all projections:
\begin{itemize}
\item Assume air resistance to be zero
\item Resolve horizontal and vertical motion
\item Horizontal - Constant speed
\item Vertical - Constant acceleration

\end{itemize}

\section{Angular projections}
The same as horizontal projections but the initial vertical velocity isn't zero.
Example:
A particle is projected at a speed of $49ms^{-1}$ at an angle of $45 \degree$ above the horizontal.
\textit{What is the time taken for the particle to reach its maximum height?}
\begin{itemize}
\item u=$49\sin45$
\item v=0
\item a=-g
\item t=?
\end{itemize}
$0=49\sin45-gt$
$t=\dfrac{49\sin45}{g}=\dfrac{5\sqrt{2}}{2}\approx3.54$
\textit{What is the maximum height reached?}
\begin{itemize}
\item u=$49\sin45$
\item v=0
\item a=-g
\item s=?
\end{itemize}
$v^2=u^2+2as$
0=$(49\sin45)^2-2gs$
$S=\dfrac{(49\sin45)^2}{2g}=61.3$
\textit{What is the time of the flight?}
\begin{itemize}
\item u=$49\sin45$
\item a=-g
\item S=0
\item t=?
\end{itemize}
$S=ut+\dfrac{1}{2}at^2$
$ $
$0=(49\sin45)t-\dfrac{1}{2}gt^2$
0=$t(49\sin45-\dfrac{gt}{2}$
$t=0$
$49\sin45=\dfrac{gt}{2}$
$t=\dfrac{2\times49\sin45}{g}=7.07$
\textit{What is the horizontal range of the particle?}
\begin{itemize}
\item t=7.07
\item Speed=$49\cos45$
\end{itemize}
$S=45\cos45\times7.07=245$
\end{obeylines}
\section{Displacement, velocity and acceleration}
$v=\dfrac{dx}{dt}$\\
\\
$a=\dfrac{dv}{dt}$\\
\\
\begin{tikzpicture}
  \node (disp) {\textbf{Displacement} $x$};
  \node[right=2cm of disp] (vel) {\textbf{Velocity} $v$};
  \node[right=2cm of vel] (acc) {\textbf{Acceleration} $a$};
  \draw[-stealth,shorten >= 2pt] (disp.north) to[bend left] node[midway,above] {Differentiate} (vel.north);
  \draw[-stealth] (vel.south) to[bend left] node[midway,below] {Integrate} (disp.south);
  \draw[-stealth,shorten >= 2pt] (vel.north) to[bend left] node[midway,above] {Differentiate} (acc.north);
    \draw[-stealth] (acc.south) to[bend left] node[midway,below] {Integrate} (vel.south);
\end{tikzpicture}
\newpage
\begin{center}
\underline{\huge Kinematics Example - Problems with calculus}
\end{center}
\textit{A particle P moves along the x-axis in a straight line so that, at time t seconds, the velocity of P is $v \ \text{ms}^{-1}$, where}\\
\\
$
  v=\left\{
  \begin{array}{@{}ll@{}}
    10t-2t^2, & \text{for}\ 0\leqslant t \leqslant 6  \\
    -\dfrac{432}{t^2}, & t>6
  \end{array}\right.
$\\
\\
\\
\textit{At $t = 0$, P is at the origin O. Find the displacement of P from O when}\\
\\
$\mathit{t = 6}$\\
\\
\textbf{Integrate velocity}\\
\\
\textcolor{red}{$\mathlarger{x=\int v \ dt}$}\\
\\
$x=\int 10t-2t^2 \ dt= 5t^2-\frac{2}{3}t^3+c$\\
\\
\\
\textbf{Substitute in the value of t}\\
\\
$t=6$\\
\\
$x=5\times6^2-\frac{2}{3}\times6^3=36m$\\
\\
\\
$\mathit{t = 10}$
\\
\textbf{Integrate velocity for the second half of the journey}\\
\\
$x=\mathlarger{\int-\dfrac{432}{t^2} \ dt=\int-432t^{-2} \ dt=\dfrac{-432t^{-1}}{-1}+k=\dfrac{432}{t}+k}$\\
\\
\textbf{Find value of k by using known distance at t=6}\\
\\
$t=6 \ \ x=36$\\
\\
$36=\dfrac{432}{6}+k$\\
\\
$k=36-72=-36$\\
\\
\textbf{Find distance using value of k and t}\\
\\
$t=10$\\
\\
$x=\dfrac{432}{10}-36=7.2\text{m}$
\newpage
\begin{center}
\underline{\huge Kinematics Example - Finding direction of motion}
\end{center}
\textit{At $t=0$ a particle P is projected from a fixed point O with velocity $(7i+7\sqrt{3})ms^{-1}$. The particle moves freely under gravity. The position vector of a point on the path of P is $(x\mathbf{i}+y\mathbf{j})$m relative to O.}\\
\textit{Show that:}
$$y=\sqrt{3}x-\frac{g}{98}x^2$$
\\
\\
$x$ has constant velocity so write in terms of t
\begin{equation}\label{eq:1}
x=7t
\end{equation}
Write an equation for y using $s=ut+\frac{1}{2}at^2$
\begin{equation}\label{eq:2}
y=7\sqrt{3}t-\frac{g}{2}t^2
\end{equation}
Substitute \eqref{eq:1} into the \eqref{eq:2}
$$y=\sqrt{3}x-\frac{g}{2}\times\Big(\frac{x}{7}\Big)^2$$
Simplify
\begin{equation}\label{eq:3}
y=\sqrt{3}x-\frac{g}{98}x^2
\end{equation}
\\
\textit{Find the direction of motion of P when it passes through the point on the path where $x=20$}\\
\\
Differentiate \eqref{eq:3}
$$\frac{dy}{dx}=\sqrt{3}-\frac{2gx}{98}$$
Substitute in the value of x
$$\frac{dy}{dx}=\sqrt{3}-\frac{40g}{98}$$
\\
\textcolor{red}{Arctan(gradient)=Angle to the positive horizontal as $\dfrac{dy}{dx}$ is the same as $\dfrac{O}{A}$
}
\\
Perform arctan on the gradient to find the angle
$$\arctan\Big(\sqrt{3}-\frac{40g}{98}\Big)=-66.2$$
\newpage
\begin{center}
\underline{\Huge Problems with Vectors}
\end{center}
\includegraphics[width=10cm]{vector_problems.png}\\
\textit{A ball B of mass 0.4 kg is struck by a bat at a point O which is 1.2 m above horizontal ground.
The unit vectors i and j are respectively horizontal and vertical. Immediately before being
struck, B has velocity $(-20i+4j) \ ms^{–1}$. Immediately after being struck it has velocity $(15i+
16j) \ ms^{–1}$.\\
After B has been struck, it moves freely under gravity and strikes the ground at the point A, as
shown in the diagram above. The ball is modelled as a particle.}\\
\\
\textit{Calculate the magnitude of the impulse exerted by the bat on B. }
\textcolor{red}{$$I=m(v-u)$$}
Calculate the impulse as a vector using information from the question
$$I=0.4(15i+16j-(-20i+4j)=14i+4.8j$$
Calculate the magnitude of this vector
$$|I|=\sqrt{14^2+4.8^2}=14.5Ns$$
\\
\textit{By using the principle of conservation of energy, or otherwise, find the speed of B when
it reaches A.}
Initial energy:
$$\frac{1}{2}\times0.4\times\Big(\sqrt{15^2+16^2}\Big)^2=96.2J$$
Energy gained from loss in GPE
$$0.4\times9.8\times1.2=4.704J$$
Add energies and set equal to final kinetic energy
$$4.704+96.2=\frac{1}{2}\times0.4\times v^2$$
Find v
$$v=\sqrt{\frac{4.704+96.2}{0.5\times0.4}}=22.46$$
\\
\textit{Calculate the angle which the velocity of B makes with the ground when B reaches A.}\\
Using the knowledge that horizontal velocity is constant and the speed is known, use cos to find the angle
$$\arccos\Bigg(\frac{15}{22.5}\Bigg)=48\degree$$
\\
\textit{State two additional physical factors which could be taken into account in a refinement of
the model of the situation which would make it more realistic}
\begin{itemize}
\item Air resistance
\item Wind
\item Rotation of ball (ball is not a particle)
\end{itemize}

>>>>

\begin{document}
\begin{center}
\underline{\huge M2 Notes}
\end{center}
\section{Centres of mass}
\subsection{Centre of mass of a discrete mass distribution}
$$\overline{x}=\frac{\sum m_ix_i}{\sum m_i}$$
\section{Collisions}
\subsection{Impulse and Momentum}
$$I=m(v-u)$$
\subsection{Coefficient of restitution}
$$e=\frac{\textrm{Speed of seperation}}{\textrm{Speed of approach}}$$
This can be rearranged to give in in terms of heights
$$e=\frac{\sqrt{h_2}}{\sqrt{h_1}}$$
\section{Kinematics}
\begin{tikzpicture}
  \node (disp) {\textbf{Displacement} $x$};
  \node[right=2cm of disp] (vel) {\textbf{Velocity} $v$};
  \node[right=2cm of vel] (acc) {\textbf{Acceleration} $a$};
  \draw[-stealth,shorten >= 2pt] (disp.north) to[bend left] node[midway,above] {Differentiate} (vel.north);
  \draw[-stealth] (vel.south) to[bend left] node[midway,below] {Integrate} (disp.south);
  \draw[-stealth,shorten >= 2pt] (vel.north) to[bend left] node[midway,above] {Differentiate} (acc.north);
    \draw[-stealth] (acc.south) to[bend left] node[midway,below] {Integrate} (vel.south);
\end{tikzpicture}
\section{Work, Energy and Power}
$$\textrm{Work}=\textrm{Force}\times\textrm{Distance}$$
$$WD=\mu R\times\textrm{Distance}$$
$$E_K=\frac{1}{2}mv^2$$
$$E_P=mgh$$
$$P=Fv$$
